---
title: "BrainAge_FeatureSelection"
output: html_document
date: "2023-03-24"
---

# Packages
```{r}
library(BioAge)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(Metrics)
library(stats)
library(skimr)
library(pcaPP)
```


# Read in Data
```{r}
healthy_cortthick_cbcl_phys <- read.csv("/Users/hansoochang/Drexel/ABCD/data/healthy_cortthick_cbcl_phys.csv")
healthy_gmv_cbcl_phys <- read.csv("/Users/hansoochang/Drexel/ABCD/data/healthy_brain_cbcl_phys.csv")

unhealthy_cortthick_cbcl_phys <- read.csv("/Users/hansoochang/Drexel/ABCD/data/unhealthy_cortthick_cbcl_phys.csv")
unhealthy_gmv_cbcl_phys <- read.csv("/Users/hansoochang/Drexel/ABCD/data/unhealthy_brain_cbcl_phys.csv")

```




# Check the relationship between CA and ROIs as a table
## IMPORTANT: Kendall's Tau takes too long in regular cor() function. Use cor.fk from pcaPP package
```{r}
cor_mat_non_para <- function(df) {
  
  df$age <- as.numeric(df$interview_age)/12
  df <- df[complete.cases(df),]
  
  cor_sub <- df %>% dplyr::select(contains(c("smri", "age")))
  cor_sub <- cor_sub %>% dplyr::select(-interview_age)
  
  smri_list <- colnames(df %>% dplyr::select(contains("smri")))
  cor_matrix <- cor.fk(cor_sub)
  correlations <- cor_matrix["age", smri_list]
  
  # Matrix with highest absolute correlations between ROI and age
  cor_mat_kendall <- sort(abs(correlations), decreasing = T)
  
  # Convert to Pearson's r and Cohen's d for effect size comparison
  cor_mat_kendall_to_r <- sin(pi*0.5*cor_mat_kendall)
  cor_mat_kendall_to_d <- 2*cor_mat_kendall_to_r/sqrt(1-cor_mat_kendall_to_r**2)
  
  return(cor_mat_kendall_to_r)
}

```
https://digitalcommons.wayne.edu/cgi/viewcontent.cgi?article=1719&context=jmasm

Cohen (1988) defined the values of effect sizes
for this group as small d = .20, medium d = .50,
and large d = .80.


# Run Custom Cor Function for Cortical Thickness and GMV
## No gender split
```{r}
cortthick_cors <- cor_mat_non_para(healthy_cortthick_cbcl_phys)
gmv_cors <- cor_mat_non_para(healthy_gmv_cbcl_phys)

cortthick_cors
```

## Gender Split (only gmv for now)
```{r}
healthy_gmv_cbcl_phys_m <- healthy_gmv_cbcl_phys %>% filter(sex.x == "M")
healthy_gmv_cbcl_phys_f <- healthy_gmv_cbcl_phys %>% filter(sex.x == "F")

unhealthy_gmv_cbcl_phys_m <- unhealthy_gmv_cbcl_phys %>% filter(sex.x == "M")
unhealthy_gmv_cbcl_phys_f <- unhealthy_gmv_cbcl_phys %>% filter(sex.x == "F")

```


```{r}
gmv_cors_m <- cor_mat_non_para(healthy_gmv_cbcl_phys_m)
gmv_cors_f <- cor_mat_non_para(healthy_gmv_cbcl_phys_f)
```


# Check Median
## No Split
```{r}
as.data.frame(cortthick_cors)
as.data.frame(gmv_cors)

median(gmv_cors)
median(cortthick_cors)
```

## Gender Split
```{r}
as.data.frame(gmv_cors_m)
as.data.frame(gmv_cors_f)

median(gmv_cors_m)
median(gmv_cors_f)
```


# Subset original df (Get any ROIs with r above .1)
## No Split
```{r}
sub_rois_cortthick <- as.data.frame(cortthick_cors) %>% filter(cortthick_cors < 0.1072957) %>% select(-cortthick_cors)

# Using Median here
sub_rois_gmv <- as.data.frame(gmv_cors) %>% filter(gmv_cors < 0.03979008) %>% select(-gmv_cors)

sub_rois_cortthick <- tibble::rownames_to_column(sub_rois_cortthick, "a") 
sub_rois_list_cortthick <- sub_rois_cortthick$a

sub_rois_gmv <- tibble::rownames_to_column(sub_rois_gmv, "a") 
sub_rois_list_gmv <- sub_rois_gmv$a

healthy_cortthick_cbcl_phys_sub <- healthy_cortthick_cbcl_phys %>% select(-sub_rois_list_cortthick)
unhealthy_cortthick_cbcl_phys_sub <- unhealthy_cortthick_cbcl_phys %>% select(-sub_rois_list_cortthick)

healthy_gmv_cbcl_phys_sub <- healthy_gmv_cbcl_phys %>% select(-sub_rois_list_gmv)
unhealthy_gmv_cbcl_phys_sub <- unhealthy_gmv_cbcl_phys %>% select(-sub_rois_list_gmv)

write.csv(healthy_cortthick_cbcl_phys_sub, "/Users/hansoochang/Drexel/ABCD/data/healthy_cortthick_cbcl_phys_sub.csv")
write.csv(unhealthy_cortthick_cbcl_phys_sub, "/Users/hansoochang/Drexel/ABCD/data/unhealthy_cortthick_cbcl_phys_sub.csv")

write.csv(healthy_gmv_cbcl_phys_sub, "/Users/hansoochang/Drexel/ABCD/data/healthy_gmv_cbcl_phys_sub.csv")
write.csv(unhealthy_gmv_cbcl_phys_sub, "/Users/hansoochang/Drexel/ABCD/data/unhealthy_gmv_cbcl_phys_sub.csv")
```


## Gender split
```{r}
# Male
sub_rois_gmv_m <- as.data.frame(gmv_cors_m) %>% filter(gmv_cors_m < median(gmv_cors_m)) %>% select(-gmv_cors_m)

sub_rois_gmv_m <- tibble::rownames_to_column(sub_rois_gmv_m, "a") 
sub_rois_list_gmv_m <- sub_rois_gmv_m$a

healthy_gmv_cbcl_phys_sub_m <- healthy_gmv_cbcl_phys_m %>% select(-sub_rois_list_gmv_m)
unhealthy_gmv_cbcl_phys_sub_m <- unhealthy_gmv_cbcl_phys_m %>% select(-sub_rois_list_gmv_m)

write.csv(healthy_gmv_cbcl_phys_sub_m, "/Users/hansoochang/Drexel/ABCD/data/healthy_gmv_cbcl_phys_sub_m.csv")
write.csv(unhealthy_gmv_cbcl_phys_sub_m, "/Users/hansoochang/Drexel/ABCD/data/unhealthy_gmv_cbcl_phys_sub_m.csv")

# Female
sub_rois_gmv_f <- as.data.frame(gmv_cors_f) %>% filter(gmv_cors_f < median(gmv_cors_f)) %>% select(-gmv_cors_f)

sub_rois_gmv_f <- tibble::rownames_to_column(sub_rois_gmv_f, "a") 
sub_rois_list_gmv_f <- sub_rois_gmv_f$a

healthy_gmv_cbcl_phys_sub_f <- healthy_gmv_cbcl_phys_f %>% select(-sub_rois_list_gmv_f)
unhealthy_gmv_cbcl_phys_sub_f <- unhealthy_gmv_cbcl_phys_f %>% select(-sub_rois_list_gmv_f)


write.csv(healthy_gmv_cbcl_phys_sub_f, "/Users/hansoochang/Drexel/ABCD/data/healthy_gmv_cbcl_phys_sub_f.csv")
write.csv(unhealthy_gmv_cbcl_phys_sub_f, "/Users/hansoochang/Drexel/ABCD/data/unhealthy_gmv_cbcl_phys_sub_f.csv")


```
























