---
title: "BrainAge_GMV"
output: html_document
date: "2023-03-15"
---


# Libraries
```{r}
library(BioAge)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(Metrics)
library(stats)
library(skimr)
library(stringr)
library(devtools)
library(neuroCombat)
library(tidyr)
library(reshape2)
```

# Read in data and get only necessary columns
```{r}
smrip <- read.delim("/Users/hansoochang/Drexel/ABCD/data/abcd_smrip10201.txt")
colnames(smrip)

# Select only cortical volume measures
# Include summary measures as well for now
# Make sure to not include white matter (ex. corpus callosum) or ventricles
smrip_gvol <- smrip %>% select(c("subjectkey", "sex", "eventname", "interview_age",
                                "smri_vol_cdk_banksstslh":"smri_vol_cdk_insularh", 
                            "smri_vol_cdk_totallh":"smri_vol_cdk_total", 
                            "smri_vol_scs_crbcortexlh":"smri_vol_scs_pallidumlh",
                            "smri_vol_scs_hpuslh":"smri_vol_scs_amygdalalh", 
                            "smri_vol_scs_aal":"smri_vol_scs_vedclh", 
                            "smri_vol_scs_crbcortexrh":"smri_vol_scs_amygdalarh", 
                            "smri_vol_scs_aar":"smri_vol_scs_vedcrh", "smri_vol_scs_subcorticalgv"))

smrip_gvol[smrip_gvol == ""] <- NA
smrip_gvol <- smrip_gvol[-1,]
smrip_gvol <- smrip_gvol[complete.cases(smrip_gvol), ]

skim(smrip_gvol)
head(smrip_gvol)
```

```{r}
smrip_h %>% filter(src_subject_id == "NDAR_INV022ZVCT8")
smrip_gvol %>% filter(subjectkey == "NDAR_INV022ZVCT8")
```


# Read in Harmonized Data
```{r}
smrip_h <- read.csv("/Users/hansoochang/Drexel/ABCD/data/sMRI_CorticalVolume-APARC_2yearfollow-up_ComBatHarmonized_interview_age.csv")
tail(smrip_h)
colnames(smrip_h)

# Select only cortical volume measures
# Include summary measures as well for now
# Make sure to not include white matter (ex. corpus callosum) or ventricles
smrip_gvol <- smrip %>% select(c("subjectkey", "sex", "eventname", "interview_age",
                                "smri_vol_cdk_banksstslh":"smri_vol_cdk_insularh", 
                            "smri_vol_cdk_totallh":"smri_vol_cdk_total", 
                            "smri_vol_scs_crbcortexlh":"smri_vol_scs_pallidumlh",
                            "smri_vol_scs_hpuslh":"smri_vol_scs_amygdalalh", 
                            "smri_vol_scs_aal":"smri_vol_scs_vedclh", 
                            "smri_vol_scs_crbcortexrh":"smri_vol_scs_amygdalarh", 
                            "smri_vol_scs_aar":"smri_vol_scs_vedcrh", "smri_vol_scs_subcorticalgv"))

smrip_gvol[smrip_gvol == ""] <- NA
smrip_gvol <- smrip_gvol[-1,]
smrip_gvol <- smrip_gvol[complete.cases(smrip_gvol), ]

skim(smrip_gvol)
head(smrip_gvol)
```




# Quality Control for sMRI (Freesurfer)
Exclusion is recommended if any of the five categories are rated as severe (a value of 3).
https://www.sciencedirect.com/science/article/pii/S1053811919306822

2.5.4. Post-processed sMRI quality control
reviewers gauge the severity of five categories of image artifact or reconstruction inaccuracy: motion, intensity inhomogeneity, white matter underestimation, pial overestimation, and magnetic susceptibility artifact. Numeric values are assigned on a scale of 0–3, indicating absent, mild, moderate, and severe levels of each type of artifact,
respectively. The reviewers assign an overall QC score indicating whether the cortical surface reconstruction is recommended for use (1) or
recommended for exclusion (0). Exclusion is recommended if any of the five categories are rated as severe (a value of 3).

```{r}
qc <- read.delim("/Users/hansoochang/Drexel/ABCD/data/abcd_fsurfqc01.txt")
qc[qc == ""] <- NA
qc <- qc[-1,]

# Remove any empty cases
qc <- subset(qc, complete.cases(qc[, c("subjectkey", "eventname")]))

# Replace 0 with baseline_year_1_arm_1 for eventname
qc$eventname <- ifelse(qc$eventname == "0", "baseline_year_1_arm_1", qc$eventname)

# Remove any subjectkey less than 16 characters long
qc$subjectkey <- ifelse(nchar(qc$subjectkey) < 16, NA, qc$subjectkey)

qc <- subset(qc, complete.cases(qc[, c("subjectkey", "eventname")]))

qc$paired <- paste(qc$subjectkey, qc$eventname)

skim(qc)
```


# Select all rows with any artifacts above 3 and any value of fsqc_qc of 1
```{r}
# any value of fsqc_qc of 1
qc_overall <- qc[qc$fsqc_qc == "1",]
qc_overall <- qc_overall[!is.na(qc_overall$collection_id),]
nrow(qc_overall) # 1671

qc_art <- qc %>% select(subjectkey, sex, eventname, paired, fsqc_nrev:fsqc_qu_artifact)

qc_art_meas <- qc_art %>% select(fsqc_qu_motion:fsqc_qu_artifact)

# df with any measure at 3 (severe quality issue)
qc_art_3 <- qc_art[apply(qc_art_meas > 2, 1, any, na.rm = TRUE), ]

nrow(qc_art_3) # 942
```


# Remove participants with rejected QC values from smrip_gvol
```{r}
nrow(smrip_gvol) #44197
nrow(qc_overall) #1671
nrow(qc_art_3) #942

# Remove any value of fsqc_qc of 1 from main GMV df
smrip_gvol_qc <- anti_join(smrip_gvol, qc_overall, by = c("subjectkey", "eventname"))
nrow(smrip_gvol_qc) # 42220

# Remove any artifacts above value of 3
smrip_gvol_qc <- anti_join(smrip_gvol_qc, qc_art_3, by = c("subjectkey", "eventname"))
nrow(smrip_gvol_qc) # 41069

smrip_dup <- smrip_gvol_qc[duplicated(smrip_gvol_qc$subjectkey) & duplicated(smrip_gvol_qc$eventname), ]

# dataframe with duplicates in the post-qc smrip df
## NB: Takes some time to run
nrow(smrip_dup[order(smrip_dup$subjectkey),]) # 29703


# Keep only one of the duplicated values.
smrip_gvol_qc_uni <- unique(smrip_dup, by = c("subjectkey", "eventname"))
nrow(smrip_gvol_qc_uni) #16499

head(smrip_gvol_qc_uni[order(smrip_gvol_qc_uni$subjectkey, smrip_gvol_qc_uni$eventname),], 100)
```

# Convert sMRI values to numeric
```{r}
smrip_gvol_qc_uni <- smrip_gvol_qc_uni %>% 
  mutate_at(vars(contains("smri")), as.numeric)

smrip_gvol_qc_uni <- smrip_gvol_qc_uni[complete.cases(smrip_gvol_qc_uni), ]

skim(smrip_gvol_qc_uni) # 16369
```


# Correct for Scanner differences 
## read in file and merge
```{r}
mri <- read.delim("/Users/hansoochang/Drexel/ABCD/data/abcd_mri01.txt")
mri[mri == ""] <- NA
mri <- mri[-1,]

# merge mri and gmv
gmv_mri <- merge(smrip_gvol_qc_uni, mri, by = c("subjectkey", "eventname"))
colnames(gmv_mri)

gmv_mri <- gmv_mri %>% select(contains(c("subjectkey", "eventname", "smri", "mri_info")))

```

## prepare data for neuroCombat
```{r}
p=10000
n=10
batch = c(1,1,1,1,1,2,2,2,2,2) #Batch variable for the scanner id
dat = matrix(runif(p*n), p, n) #Random Data matrix
dat
data.harmonized <- neuroCombat(dat=dat, batch=batch)

head(gmv_mri)

nrow(gmv_mri) # 35979
gmv_mri <- gmv_mri %>% filter(grepl(paste0("^", "HASH"), mri_info_deviceserialnumber) &
                        nchar(mri_info_deviceserialnumber) >= 10 & nchar(mri_info_deviceserialnumber) <= 20)
nrow(gmv_mri) # 35890

# 33 unique scanners
table(gmv_mri$mri_info_deviceserialnumber)

gmv_mri$paste <- paste(gmv_mri$subjectkey, gmv_mri$eventname)

dupes <- duplicated(gmv_mri)
gmv_mri_unique <- gmv_mri[!dupes, ]
gmv_mri_unique <- gmv_mri_unique[!duplicated(gmv_mri_unique$paste),]
```


```{r}
head(gmv_mri_unique)
gmv_mri_unique_y2 <- gmv_mri_unique %>% filter(eventname == "2_year_follow_up_y_arm_1")
gmv_mri_unique_b <- gmv_mri_unique %>% filter(eventname == "baseline_year_1_arm_1")


melted_df <- melt(gmv_mri_unique_y2, id.vars = c("subjectkey", "eventname"))
wide_df <- dcast(melted_df, variable + eventname ~ subjectkey, value.var = "value")


match_list <- colnames(gmv_mri_unique_y2 %>% select(contains(c("smri", "mri_info_deviceserialnumber"))))

gmv_mri_dat <- wide_df %>% filter(variable %in% match_list) %>% 
  select(contains(c("variable", "NDAR"))) 

gmv_mri_dat <- gmv_mri_dat[complete.cases(gmv_mri_dat), ]

gmv_mri_dat[-91, -1] <- apply(gmv_mri_dat[-91, -1], 2, as.numeric)

gmv_mri_dat <- gmv_mri_dat %>% select_if(~all(. >= 0))

gmv_mri_batch <- gmv_mri_dat[91,]
gmv_mri_dat <- gmv_mri_dat[-91,]

gmv_mri_dat <- mutate_all(gmv_mri_dat, function(x) as.numeric(as.character(x)))


length(gmv_mri_batch)
dim(gmv_mri_dat)

```

## run neuroCombat
```{r}
rownames(gmv_mri_dat) <- NULL
colnames(gmv_mri_dat) <- NULL


as.character(gmv_mri_batch[1,])

data.harmonized <- neuroCombat(dat=data.matrix(gmv_mri_dat), batch=as.character(gmv_mri_batch), eb = FALSE, parametric = FALSE)
View(data.harmonized$dat.combat)
View(gmv_mri_dat)

```



# Check Distributions.
There seem to be outliers
```{r}
smri_list <- colnames(smrip_gvol_qc_uni %>% select(contains("smri")))

for (smri in 1:length(smri_list)) {
  hist(smrip_gvol_qc_uni[, smri_list[smri]], xlab = smri_list[smri])
}

```

# Remove Impossible Values
```{r}
# filter for any values greater than 3 SDs from the mean by column
smri_outlier <- smrip_gvol_qc_uni %>% mutate(across(contains("smri"), ~ abs(scale(.)) > 3)) %>%
  filter(rowSums(across(contains("smri"))) > 0)

smri_outlier$paired <- paste(smri_outlier$subjectkey, smri_outlier$eventname)

smrip_gvol_qc_uni$paired <- paste(smrip_gvol_qc_uni$subjectkey, smrip_gvol_qc_uni$eventname)

# These are major outlier. Impossible values.
smrip_gvol_qc_uni_outlier <- smrip_gvol_qc_uni %>% filter(paired %in% smri_outlier$paired)

smrip_gvol_qc_uni_norm <- smrip_gvol_qc_uni %>% filter(!(paired %in% smri_outlier$paired))

skim(smrip_gvol_qc_uni_outlier)
skim(smrip_gvol_qc_uni_norm)


### Check why there are so many negative values and very small values
sort(smrip_gvol_qc_uni$smri_vol_cdk_banksstslh)
sort(smrip_gvol_qc_uni$smri_vol_cdk_cdacatelh)
```


# See Distributions of truncated data
```{r}
smri_list <- colnames(smrip_gvol_qc_uni_norm %>% select(contains("smri")))

for (smri in 1:length(smri_list)) {
  hist(smrip_gvol_qc_uni_norm[, smri_list[smri]], xlab = smri_list[smri])
}

skim(smrip_gvol_qc_uni_norm)
```
Check smri_vol_cdk_sufrrh to smri_vol_scs_subcroticalgv

# Remove more extreme outliers
```{r}
more_outliers <- smrip_gvol_qc_uni_norm %>% select(paired, smri_vol_cdk_sufrrh:smri_vol_scs_subcorticalgv)
more_outliers <- more_outliers %>% filter_all(any_vars(. < 100))

smri_big_outlier <- more_outliers %>% mutate(across(contains("smri"), ~ abs(scale(.)) > 3)) %>%
  filter(rowSums(across(contains("smri"))) > 0)

# Remove any value below 100 which would be impossible
smrip_gvol_qc_uni_big_norm <- 
  smrip_gvol_qc_uni_norm[!apply(smrip_gvol_qc_uni_norm %>% 
                                    select("smri_vol_cdk_banksstslh":"smri_vol_scs_subcorticalgv") < 100, 
                                1, any, na.rm=TRUE), ]

```


# Check Distributions of all outliers removed
```{r}
smri_list <- colnames(smrip_gvol_qc_uni_big_norm %>% select(contains("smri")))

for (smri in 1:length(smri_list)) {
  hist(smrip_gvol_qc_uni_big_norm[, smri_list[smri]], xlab = smri_list[smri])
}

skim(smrip_gvol_qc_uni_big_norm)

```
Seems pretty good. Will ask about large values.

# Remove more outliers extremes
```{r}
smrip_gvol_qc_uni_big_norm <- smrip_gvol_qc_uni_big_norm %>% filter(!(smri_vol_cdk_lobfrlh < 5000 |
                                      smri_vol_cdk_mobfrlh < 3000 |
                                      smri_vol_cdk_paracnlh > 8000 |
                                      smri_vol_cdk_parsobislh < 1500 |
                                      smri_vol_cdk_postcnlh > 22500 |
                                      smri_vol_cdk_precnlh < 7500 | 
                                      smri_vol_cdk_rrmdfrlh < 7500 | 
                                      smri_vol_cdk_insulalh > 12000 | smri_vol_cdk_insulalh < 3000 |
                                      smri_vol_cdk_fusiformrh < 5000 |
                                      smri_vol_cdk_ifplrh < 6000 |
                                      smri_vol_cdk_iftmrh < 6000 |
                                      smri_vol_cdk_ihcaterh < 1000 | smri_vol_cdk_ihcaterh > 5500 |
                                      smri_vol_cdk_paracnrh > 9000 |
                                      smri_vol_cdk_totallh < 1e+05 |
                                      smri_vol_cdk_totalrh < 1e+05 | 
                                      smri_vol_scs_crbcortexlh < 2000 | 
                                      smri_vol_scs_tplh < 2000 | smri_vol_scs_tplh > 15000 |
                                      smri_vol_scs_caudatelh < 1000 | smri_vol_scs_caudatelh > 7500 |
                                      smri_vol_scs_putamenlh < 1000 | smri_vol_scs_putamenlh > 7500 |
                                      smri_vol_scs_pallidumlh > 5000 |
                                      smri_vol_scs_hpuslh > 20000 | 
                                      smri_vol_scs_amygdalalh > 5000 | 
                                      smri_vol_scs_aal > 8000 | 
                                      smri_vol_scs_vedclh < 1000 | smri_vol_scs_vedclh > 10000 | 
                                      smri_vol_scs_crbcortexrh < 20000 | smri_vol_scs_crbcortexrh < 30000 | 
                                      smri_vol_scs_pallidumrh > 3250 |
                                      smri_vol_scs_vedcrh < 2000 | smri_vol_scs_vedcrh > 7000 | 
                                      smri_vol_scs_hpuslh > 7000 |
                                      smri_vol_scs_subcorticalgv < 30000 | smri_vol_scs_subcorticalgv > 100000))

nrow(smrip_gvol_qc_uni_big_norm) # 14558

```

# Check Distributions again
```{r}
smri_list <- colnames(smrip_gvol_qc_uni_big_norm %>% select(contains("smri")))

for (smri in 1:length(smri_list)) {
  hist(smrip_gvol_qc_uni_big_norm[, smri_list[smri]], xlab = smri_list[smri])
}

skim(smrip_gvol_qc_uni_big_norm)
```

# Split into Timepoints (Without removing Outliers)
## We will remove outliers after splitting into healthy and unhealthy
## For comparison with Bio Age, we only need year 2
```{r}
smrip_gvol_qc_uni_y2 <- smrip_gvol_qc_uni %>% filter(eventname == "2_year_follow_up_y_arm_1")
smrip_gvol_qc_uni_base <- smrip_gvol_qc_uni %>% filter(eventname == "baseline_year_1_arm_1")

nrow(smrip_gvol_qc_uni_y2) # 6651
nrow(smrip_gvol_qc_uni_base) # 9721

write.csv(smrip_gvol_qc_uni, "/Users/hansoochang/Drexel/ABCD/data/smrip_gmv.csv")
write.csv(smrip_gvol_qc_uni_y2, "/Users/hansoochang/Drexel/ABCD/data/smrip_gmv_y2.csv")
write.csv(smrip_gvol_qc_uni_base, "/Users/hansoochang/Drexel/ABCD/data/smrip_gmv_base.csv")
```



# Split into Timepoints
## For comparison with Bio Age, we only need year 2
```{r}
smrip_gvol_qc_uni_y2 <- smrip_gvol_qc_uni_big_norm %>% filter(eventname == "2_year_follow_up_y_arm_1")
smrip_gvol_qc_uni_base <- smrip_gvol_qc_uni_big_norm %>% filter(eventname == "baseline_year_1_arm_1")

nrow(smrip_gvol_qc_uni_y2) # 5977
nrow(smrip_gvol_qc_uni_base) # 8581

write.csv(smrip_gvol_qc_uni_big_norm, "/Users/hansoochang/Drexel/ABCD/data/smrip_gmv.csv")
write.csv(smrip_gvol_qc_uni_y2, "/Users/hansoochang/Drexel/ABCD/data/smrip_gmv_y2.csv")
write.csv(smrip_gvol_qc_uni_base, "/Users/hansoochang/Drexel/ABCD/data/smrip_gmv_base.csv")
```




# Get Scanner Info
```{r}

```













