---
title: "BrainAge_Initial"
output: pdf_document
date: "2023-03-03"
---

# Libraries
```{r}
library(BioAge)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(Metrics)
library(stats)
library(skimr)
```




# Read in data and get only necessary columns
```{r}
mrisdp <- read.delim("/Users/hansoochang/Drexel/ABCD/data/abcd_mrisdp10201.txt")

# Select only cortical volume measures
dest_vol <- mrisdp %>% select(c("subjectkey", "sex", "eventname", "interview_age",
                                "mrisdp_454":"mrisdp_604"))
dest_vol[dest_vol == ""] <- NA
dest_vol <- dest_vol[-1,]
dest_vol <- dest_vol[complete.cases(dest_vol), ]

skim(dest_vol)
head(dest_vol)
```
Very good complete rate!

# Quality Control for sMRI
Exclusion is recommended if any of the five categories are rated as severe (a value of 3).
https://www.sciencedirect.com/science/article/pii/S1053811919306822

```{r}
qc <- read.delim("/Users/hansoochang/Drexel/ABCD/data/abcd_t2wqc01.txt")
qc[qc == ""] <- NA
qc <- qc[-1,]

table(qc$t2w_postqc_revdisp)
qc

skim(qc)
```

Find and Remove any instances with qc score of 3
```{r}
colnames(qc)

qc_sel <- qc %>% select(c("subjectkey", "abcd_t2wqc01_id", "eventname", 
                "t2w_postqc_containertype":"t2w_postqc_notes"))
qc_sel_meas <- qc_sel %>% select(t2w_postqc_containertype:t2w_postqc_notes)

qc_sel_3 <- qc_sel[apply(qc_sel_meas > 2, 1, any, na.rm = TRUE), ]

# df with any measure at 3 (severe quality issue)
qc_sel_3


```

# Merge dataframes and find any rows with 3 in any qc score
```{r}
brain_qc_merged <- merge(dest_vol, qc_sel, by = c("subjectkey", "eventname"))
brain_qc_merged_meas <- brain_qc_merged %>% select(t2w_postqc_containertype:t2w_postqc_notes)
brain_qc_merged_g3 <- brain_qc_merged[apply(brain_qc_merged_meas > 2, 1, any, na.rm = TRUE), ]

brain_qc_merged_g3
```

# Get rows with qc values less than 3
## the dataframes below will bbe okay by qc standards
```{r}
anti_dest_vol <- anti_join(brain_qc_merged, brain_qc_merged_g3, by = c("subjectkey", "eventname"))

anti_dest_vol_markers <- colnames(anti_dest_vol %>% select(mrisdp_454:t2w_postqc_artifact))
anti_dest_vol[, anti_dest_vol_markers] <- 
  as.data.frame(apply(anti_dest_vol[, anti_dest_vol_markers], 2, as.numeric))

exact_dup_rows <- duplicated(anti_dest_vol)
nrow(anti_dest_vol[exact_dup_rows,]) #6820

dup_rows <- duplicated(anti_dest_vol %>% select("subjectkey", "eventname"))
nrow(anti_dest_vol[dup_rows,]) # 7054

# Rows with duplicate subject and time, but different values, no explanation
dup_rows_not_exact <- anti_join(anti_dest_vol[dup_rows,], anti_dest_vol[exact_dup_rows,])
```

# Merge exact duplicates into one
```{r}
anti_dest_vol <- anti_dest_vol[-which(exact_dup_rows),]

```

# Remove unexplained duplicates with different values and
```{r}
nrow(anti_dest_vol) # 18740
nrow(anti_join(anti_dest_vol, dup_rows_not_exact)) # 18546

dest_vol_no_dup_comp <- anti_join(anti_dest_vol, dup_rows_not_exact)

# Check to see if no duplicated
dest_vol_no_dup_comp[duplicated(dest_vol_no_dup_comp),]
```


# Standardization
## should be standardized by males and females from entire dataset
```{r}
# Vector of Biomarkers (brain regions)
markers <- colnames(dest_vol_no_dup_comp %>% select(mrisdp_454:mrisdp_604))

## Men (14434) ##
men_merge <- dest_vol_no_dup_comp[which(dest_vol_no_dup_comp$sex == "M"),]
nrow(men_merge) 

# scale biomarkers
men_merge[, markers] <- as.data.frame(apply(men_merge[, markers], 2, as.numeric))
men_merge[, markers] <- scale(men_merge[, markers])
## Women (12882) ##
women_merge <- dest_vol_no_dup_comp[which(dest_vol_no_dup_comp$sex == "F"),]
nrow(women_merge) 


# scale biomarkers
women_merge[, markers] <- as.data.frame(apply(women_merge[, markers], 2, as.numeric))
women_merge[, markers] <- scale(women_merge[, markers])

## Remerge men and women ##
stand_brain_dest_vol_no_dup_comp <- rbind(men_merge, women_merge)

stand_dest_vol <- stand_brain_dest_vol_no_dup_comp %>% select(subjectkey:mrisdp_604)
stand_dest_vol <- stand_dest_vol[complete.cases(stand_dest_vol), ]

head(stand_dest_vol)
dim(stand_dest_vol)

write.csv(stand_dest_vol, "/Users/hansoochang/Drexel/ABCD/data/stand_dest_vol.csv")
```

There are many duplicates. Let try to see the quality control to see if this while clarify which value to use



# Split into Timepoints
## For comparison with Bio Age, we only need year 2
```{r}
stand_dest_vol_y2 <- stand_dest_vol %>% filter(eventname == "2_year_follow_up_y_arm_1")
stand_dest_vol_base <- stand_dest_vol %>% filter(eventname == "baseline_year_1_arm_1")

nrow(stand_dest_vol_y2) # 7442
nrow(stand_dest_vol_base) # 11073

write.csv(stand_dest_vol_y2, "/Users/hansoochang/Drexel/ABCD/data/stand_dest_vol_y2.csv")
write.csv(stand_dest_vol_base, "/Users/hansoochang/Drexel/ABCD/data/stand_dest_vol_base.csv")
```





```{r}
write.csv(stand_brain_dest_vol_y2, "/Users/hansoochang/Drexel/ABCD/data/stand_brain_dest_vol_y2.csv")
write.csv(stand_brain_dest_vol_base, "/Users/hansoochang/Drexel/ABCD/data/stand_brain_dest_vol_base.csv")
```





# Descriptives
```{r}
length(unique(stand_brain_dest_vol$subjectkey)) # 11804
table(stand_brain_dest_vol$eventname)


```





