---
title: "BrainAge_KDM_V3_GMV_FSelection"
output: html_document
date: "2023-03-23"
---
```{r}
#install.packages("flexsurv")
#devtools::install_github("dayoonkwon/BioAge")
library(BioAge)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(Metrics)
library(stats)
library(skimr)
library(pcaPP)
library(stringr)
library(stats)
library(misty)
library(emmeans)
install.packages("BioStat")
library(biostat)
```


Since KDM requires chronological age, we need chronological age in our biological markers. Convert age in months to age in years. 
Files from BrainAge_KDM_V2.Rmd: 

# Split into and test set from healthy cohort
```{r}
healthy_brain_cbcl_phys_sub <- read.csv("/Users/hansoochang/Drexel/ABCD/data/healthy_gmv_cbcl_phys_sub.csv")
unhealthy_brain_cbcl_phys_sub <- read.csv("/Users/hansoochang/Drexel/ABCD/data/unhealthy_gmv_cbcl_phys_sub.csv")



colnames(healthy_brain_cbcl_phys_sub)

healthy_brain_cbcl_phys_sub$age <- as.numeric(healthy_brain_cbcl_phys_sub$interview_age)/12
healthy_brain_cbcl_phys_sub$gender <- healthy_brain_cbcl_phys_sub$sex.x

unhealthy_brain_cbcl_phys_sub$age <- as.numeric(unhealthy_brain_cbcl_phys_sub$interview_age)/12 
unhealthy_brain_cbcl_phys_sub$gender <- unhealthy_brain_cbcl_phys_sub$sex.x

# about 4 participants with no age info
healthy_brain_cbcl_phys_sub <- healthy_brain_cbcl_phys_sub[complete.cases(healthy_brain_cbcl_phys_sub), ]
unhealthy_brain_cbcl_phys_sub <- unhealthy_brain_cbcl_phys_sub[complete.cases(unhealthy_brain_cbcl_phys_sub), ]

nrow(healthy_brain_cbcl_phys_sub) # 2626
nrow(unhealthy_brain_cbcl_phys_sub) # 1054
```


 


# Train
```{r}

set.seed(18)
#bm_healthy <- merge(bm, healthy_cbcl_phys, by = "subjectkey", all = FALSE)
#nrow(bm_healthy) # 375
#bm_unhealthy <- merge(bm, unhealthy_cbcl, by = "subjectkey", all = FALSE)
#nrow(bm_unhealthy) # 283

### Healthy Training ###
samp <- sample(c(TRUE, FALSE), nrow(healthy_brain_cbcl_phys_sub), replace=TRUE, prob=c(0.6,0.4))
brain_train_sub <- healthy_brain_cbcl_phys_sub[samp, ]
brain_test_sub <- healthy_brain_cbcl_phys_sub[!samp, ]

nrow(brain_train_sub) # 1571
nrow(brain_test_sub) # 1055

```


```{r}
brain_markers_sub <- colnames(brain_train_sub %>% dplyr::select(contains("smri")))
brain_markers_sub
# kdm_train_male <- kdm_calc(male_train, biomarkers = markers)
# kdm_train_female <- kdm_calc(female_train, biomarkers = markers)'

kdm_train_brain_sub <- kdm_calc(brain_train_sub, biomarkers = brain_markers_sub)
```


#Correlation tests between kdm_age and chronological age
```{r}
cor.test(as.numeric(kdm_train_brain_sub$data$kdm), as.numeric(kdm_train_brain_sub$data$age))
rmse(as.numeric(kdm_train_brain_sub$data$kdm), as.numeric(kdm_train_brain_sub$data$age))
mae(as.numeric(kdm_train_brain_sub$data$kdm), as.numeric(kdm_train_brain_sub$data$age))

# Female
# cor.test(as.numeric(kdm_train_female$data$kdm), as.numeric(kdm_train_female$data$age))
# rmse(as.numeric(kdm_train_female$data$kdm), as.numeric(kdm_train_female$data$age))
# mae(as.numeric(kdm_train_female$data$kdm), as.numeric(kdm_train_female$data$age))


#plot age vs bioage
#plot_ba(kdm_train$data, "kdm", "KDM\nBiological Age")

```


# Project to test
```{r}
# 
kdm_test_brain_sub <- kdm_calc(brain_test_sub, biomarkers = brain_markers_sub, fit = kdm_train_brain_sub$fit, s_ba2 = 
                       kdm_train_brain_sub$fit$s_ba2)

# Female
# kdm_female <- kdm_calc(female_test, biomarkers = markers, fit = kdm_train_female$fit, s_ba2 = 
#                        kdm_train_female$fit$s_ba2)


# kdm_data <- rbind(kdm_male$data, kdm_female$data)
kdm_data_brain_sub <- kdm_test_brain_sub$data
```

# Correlation tests between kdm_age (both genders) and chronological age
```{r}
cor.test(as.numeric(kdm_data_brain_sub$kdm), as.numeric(kdm_data_brain_sub$age))
rmse(as.numeric(kdm_data_brain_sub$kdm), as.numeric(kdm_data_brain_sub$age))
mae(as.numeric(kdm_data_brain_sub$kdm), as.numeric(kdm_data_brain_sub$age))


#plot age vs bioage
plot_ba(kdm_data_brain_sub, "kdm", "KDM\nBiological Age")

```


# Project trained model onto Unhealthy cohort
## Get prediction
```{r}
# project to unhealthy Group
# males_unhealthy <- bm_unhealthy[which(bm_unhealthy$gender == "M"),]

kdm_unhealthy_brain_sub <- kdm_calc(unhealthy_brain_cbcl_phys_sub, biomarkers = brain_markers_sub, fit = kdm_train_brain_sub$fit, s_ba2 = 
                       kdm_train_brain_sub$fit$s_ba2)

# Females
#females_unhealthy <- bm_unhealthy[which(bm_unhealthy$gender == "F"),]

#kdm_female_unhealthy <- kdm_calc(females_unhealthy, biomarkers = markers, fit = #kdm_train_female$fit, s_ba2 = 
#                      kdm_train_female$fit$s_ba2)


# merge
# kdm_data_unhealthy <- rbind(kdm_male_unhealthy$data, kdm_female_unhealthy$data)
kdm_data_unhealthy_brain_sub <- kdm_unhealthy_brain_sub$data 

```


# Correlation test between KDM age and Chronological Age for Unhealthy cohort
```{r}
cor.test(as.numeric(kdm_data_unhealthy_brain_sub$kdm), as.numeric(kdm_data_unhealthy_brain_sub$age))
rmse(as.numeric(kdm_data_unhealthy_brain_sub$kdm), as.numeric(kdm_data_unhealthy_brain_sub$age))
mae(as.numeric(kdm_data_unhealthy_brain_sub$kdm), as.numeric(kdm_data_unhealthy_brain_sub$age))

```


```{r}
plot_ba(kdm_data_unhealthy_brain_sub, "kdm", "KDM\nBiological Age")
plot(x = kdm_data_unhealthy_brain_sub$age, y = kdm_data_unhealthy_brain_sub$kdm_advance)
```


# Test for Normality for KDM
```{r}
# variances are approximately = 1

# KDM Biological Age
shapiro.test(kdm_data_brain_sub$kdm)
shapiro.test(kdm_data_unhealthy_brain_sub$kdm)

ggdensity(kdm_data_brain_sub$kdm)
ggdensity(kdm_data_unhealthy_brain_sub$kdm)
ggdensity(kdm_data_brain_sub$age) # For some reason, less 12 year olds in the data

ggqqplot(kdm_data_brain_sub$kdm)
ggqqplot(kdm_data_unhealthy_brain_sub$kdm)

hist(kdm_data_brain_sub$kdm, xlab = "KDM Value")
hist(kdm_data_brain_sub$kdm_advance, xlab = "KDM Advance Value")

hist(kdm_data_unhealthy_brain_sub$kdm, xlab = "KDM Value")
hist(kdm_data_unhealthy_brain_sub$kdm_advance, xlab = "KDM Advance Value")
```



# T-Test
```{r}

sd(kdm_data_brain_sub$kdm)
sd(kdm_data_unhealthy_brain_sub$kdm)

sd(kdm_data_brain_sub$kdm_advance)
sd(kdm_data_unhealthy_brain_sub$kdm_advance)

# T-Test for average healthy KDM and unhealthy KDM
t.test(kdm_data_brain_sub$kdm, kdm_data_unhealthy_brain_sub$kdm, alternative = "two.sided", var.equal = T)

# T-Test for average healthy KDM_advance and unhealthy KDM_advance
t.test(kdm_data_brain_sub$kdm_advance, kdm_data_unhealthy_brain_sub$kdm_advance, alternative = "two.sided", var.equal = T)

# t.test between chronological ages
t.test(kdm_data_brain_sub$age, kdm_data_unhealthy_brain_sub$age, alternative = "two.sided", var.equal = TRUE)

```

# Add scanner information for covariate analyses
```{r}
mri <- read.delim("/Users/hansoochang/Drexel/ABCD/data/abcd_mri01.txt")
mri <- mri[-1,]

mri_sel <- mri %>% select(subjectkey, eventname, mri_info_manufacturer, mri_info_deviceserialnumber)

mri_sel[mri_sel == ""] <- NA

skim(mri_sel)
```
## Cleaning dataframe
```{r}
mri_sel_comp <- mri_sel[complete.cases(mri_sel),]

mri_sel_comp_1 <- mri_sel_comp %>% filter(nchar(mri_info_deviceserialnumber) == 12)

mri_sel_comp_2 <- mri_sel_comp_1 %>% filter(str_detect(mri_info_deviceserialnumber, "HASH"))

string_vec <- c("SIEMENS", "Philips Medical Systems", "GE MEDICAL SYSTEMS")

mri_sel_comp_3 <- mri_sel_comp_2 %>% filter(mri_info_manufacturer %in% string_vec)

mri_sel_comp_3

```

## Add Site
```{r}
SiteId <- read.csv("/Users/hansoochang/Drexel/ABCD/data/abcd_lt01.csv")
SiteId <- SiteId[-1,]
SiteId[SiteId == ""] <- NA

SiteId_site <- SiteId %>% select(subjectkey, eventname, site_id_l)
```



## Merge with KDM_V3 
```{r}
nrow(kdm_data_brain_sub) # 1058
nrow(kdm_data_unhealthy_brain_sub) # 1059

kdm_data_brain_sub$eventname <- kdm_data_brain_sub$eventname.x
kdm_data_unhealthy_brain_sub$eventname <- kdm_data_unhealthy_brain_sub$eventname.x

# Add Scanner
gmv_mri_healthy <- merge(kdm_data_brain_sub, mri_sel_comp_2, by = c("subjectkey", "eventname"))
gmv_mri_unhealthy <- merge(kdm_data_unhealthy_brain_sub, mri_sel_comp_2, by = c("subjectkey", "eventname"))

# Add Site ID
gmv_mri_healthy <- merge(gmv_mri_healthy, SiteId_site, by = c("subjectkey", "eventname"))
gmv_mri_unhealthy <- merge(gmv_mri_unhealthy, SiteId_site, by = c("subjectkey", "eventname"))

# Both health and unhealthy
gmv_year2_cs <- rbind(kdm_data_brain_sub, kdm_data_unhealthy_brain_sub)
dim(gmv_year2_cs)

# This file is the 1058 healthy test set participants plus the 
# 1059 unhealthy participants for feature selected gmv. We will use this to find correlations with
# CBCL, Mania, etc.
write.csv(gmv_year2_cs, "/Users/hansoochang/Drexel/ABCD/data/gmv_year2_cs.csv")


```

## Boxplots for ROIs
```{r}
table(gmv_mri_healthy$mri_info_deviceserialnumber)
table(gmv_mri_healthy$mri_info_manufacturer)

smri_list <- colnames(gmv_mri_healthy %>% select(contains("smri")))

for (smri in smri_list) {
  print(ggplot(gmv_mri_healthy, aes_string(x = "mri_info_manufacturer", y = smri)) +
  geom_boxplot())
}


```

## Boxplots for KDM
```{r}
ggplot(gmv_mri_healthy, aes(x = mri_info_manufacturer, y = kdm_advance)) +
  geom_boxplot()

ggplot(gmv_mri_unhealthy, aes(x = mri_info_manufacturer, y = kdm_advance)) +
  geom_boxplot()

ggplot(gmv_mri_healthy, aes(x = mri_info_manufacturer, y = age)) +
  geom_boxplot()

ggplot(gmv_mri_unhealthy, aes(x = mri_info_manufacturer, y = age)) +
  geom_boxplot()

```

```{r}
gmv_mri_healthy$mri_info_manufacturer <- as.factor(gmv_mri_healthy$mri_info_manufacturer)
typeof(gmv_mri_healthy$kdm_advance)

# summary(lm(data = gmv_mri_healthy, kdm_advance ~ age + mri_info_manufacturer))
# summary(lm(data = gmv_mri_unhealthy, kdm_advance ~ age + mri_info_manufacturer))

# test.welch(age ~ mri_info_manufacturer, data = gmv_mri_healthy)
# test.welch(age ~ mri_info_manufacturer, data = gmv_mri_unhealthy)

welch_anova_h <- test.welch(kdm_advance ~ mri_info_manufacturer, data = gmv_mri_healthy)
welch_anova_u <- test.welch(kdm_advance ~ mri_info_manufacturer, data = gmv_mri_unhealthy)

welch_anova_h

fit_h <- aov(kdm_advance ~ mri_info_manufacturer, data = gmv_mri_healthy)
fit_u <- aov(kdm_advance ~ mri_info_manufacturer, data = gmv_mri_unhealthy)

TukeyHSD(fit_h)
TukeyHSD(fit_u)
```



# Participants for Outcome Correlations
```{r}
nrow(kdm_data_brain_sub) # 1058
nrow(kdm_data_unhealthy_brain_sub) # 1059

gmv_year2_cs <- rbind(kdm_data_brain_sub, kdm_data_unhealthy_brain_sub)

gmv_year2_cs <- merge(gmv_year2_cs, site, by = c("subjectkey", "eventname"))



# This file is the 1058 healthy test set participants plus the 
# 1059 unhealthy participants for feature selected gmv. We will use this to find correlations with
# CBCL, Mania, etc.
write.csv(gmv_year2_cs, "/Users/hansoochang/Drexel/ABCD/data/gmv_year2_cs.csv")

```










