---
title: "BrainAge_CortThick"
output: html_document
date: "2023-03-15"
---


# Libraries
```{r}
library(BioAge)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(Metrics)
library(stats)
library(skimr)
library(stringr)
```

# Read in data and get only necessary columns
```{r}
smrip <- read.delim("/Users/hansoochang/Drexel/ABCD/data/abcd_smrip10201.txt")
colnames(smrip)

# Select only cortical Thickness measures
# Include summary measures as well for now
# Make sure to not include white matter (ex. corpus callosum) or ventricles
smrip_cortthick <- smrip %>% select(c("subjectkey", "sex", "eventname", "interview_age",
                                "smri_thick_cdk_banksstslh":"smri_thick_cdk_mean"))

smrip_cortthick[smrip_cortthick == ""] <- NA
smrip_cortthick <- smrip_cortthick[-1,]
smrip_cortthick <- smrip_cortthick[complete.cases(smrip_cortthick), ]

skim(smrip_cortthick)
head(smrip_cortthick)
```

# Quality Control for sMRI (Freesurfer)
Exclusion is recommended if any of the five categories are rated as severe (a value of 3).
https://www.sciencedirect.com/science/article/pii/S1053811919306822

2.5.4. Post-processed sMRI quality control
reviewers gauge the severity of five categories of image artifact or reconstruction inaccuracy: motion, intensity inhomogeneity, white matter underestimation, pial overestimation, and magnetic susceptibility artifact. Numeric values are assigned on a scale of 0–3, indicating absent, mild, moderate, and severe levels of each type of artifact,
respectively. The reviewers assign an overall QC score indicating whether the cortical surface reconstruction is recommended for use (1) or
recommended for exclusion (0). Exclusion is recommended if any of the five categories are rated as severe (a value of 3).

```{r}
qc <- read.delim("/Users/hansoochang/Drexel/ABCD/data/abcd_fsurfqc01.txt")
qc[qc == ""] <- NA
qc <- qc[-1,]

# Remove any empty cases
qc <- subset(qc, complete.cases(qc[, c("subjectkey", "eventname")]))

# Replace 0 with baseline_year_1_arm_1 for eventname
qc$eventname <- ifelse(qc$eventname == "0", "baseline_year_1_arm_1", qc$eventname)

# Remove any subjectkey less than 16 characters long
qc$subjectkey <- ifelse(nchar(qc$subjectkey) < 16, NA, qc$subjectkey)

qc <- subset(qc, complete.cases(qc[, c("subjectkey", "eventname")]))

qc$paired <- paste(qc$subjectkey, qc$eventname)

skim(qc)
```


# Select all rows with any artifacts above 3
```{r}
# any value of fsqc_qc of 1
qc_overall <- qc[qc$fsqc_qc == "1",]
qc_overall <- qc_overall[!is.na(qc_overall$collection_id),]
nrow(qc_overall) # 1671

qc_art <- qc %>% select(subjectkey, sex, eventname, paired, fsqc_nrev:fsqc_qu_artifact)

qc_art_meas <- qc_art %>% select(fsqc_qu_motion:fsqc_qu_artifact)

# df with any measure at 3 (severe quality issue)
qc_art_3 <- qc_art[apply(qc_art_meas > 2, 1, any, na.rm = TRUE), ]

nrow(qc_art_3) # 942
```


# Remove participants with rejected QC values from smrip_cortthick
```{r}
nrow(smrip_cortthick) #46072
nrow(qc_overall) # 1671
nrow(qc_art_3) #942

# Remove any value of fsqc_qc of 1 from main cortthick df
smrip_cortthick_qc <- anti_join(smrip_cortthick, qc_overall, by = c("subjectkey", "eventname"))
nrow(smrip_cortthick_qc) # 44037

# Remove any artifacts above value of 3
smrip_cortthick_qc <- anti_join(smrip_cortthick_qc, qc_art_3, by = c("subjectkey", "eventname"))
nrow(smrip_cortthick_qc) # 42856

# Filter duplicated rows
smrip_cortthick_dup <- smrip_cortthick_qc[duplicated(smrip_cortthick_qc$subjectkey) & duplicated(smrip_cortthick_qc$eventname), ]

# Keep only one of the duplicated values.
smrip_cortthick_qc_uni <- unique(smrip_cortthick_dup, by = c("subjectkey", "eventname"))

tail(smrip_cortthick_qc_uni[order(smrip_cortthick_qc_uni$subjectkey, smrip_cortthick_qc_uni$eventname),], 100)
```

# Convert sMRI values to numeric
```{r}
smrip_cortthick_qc_uni <- smrip_cortthick_qc_uni %>% 
  mutate_at(vars(contains("smri")), as.numeric)

smrip_cortthick_qc_uni <- smrip_cortthick_qc_uni[complete.cases(smrip_cortthick_qc_uni), ]

skim(smrip_cortthick_qc_uni)

tail(smrip_cortthick_qc_uni[order(smrip_cortthick_qc_uni$subjectkey, smrip_cortthick_qc_uni$eventname),], 100)
```

# Check Distributions.
There seem to be outliers
```{r}
smri_list <- colnames(smrip_cortthick_qc_uni %>% select(contains("smri")))

for (smri in 1:length(smri_list)) {
  hist(smrip_cortthick_qc_uni[, smri_list[smri]], xlab = smri_list[smri])
}


```

# Remove Impossible Values
```{r}
# filter for any values greater than 3 SDs from the mean by column
smri_outlier <- smrip_cortthick_qc_uni %>% mutate(across(contains("smri"), ~ abs(scale(.)) > 3)) %>%
  filter(rowSums(across(contains("smri"))) > 0)

smri_outlier$paired <- paste(smri_outlier$subjectkey, smri_outlier$eventname)

smrip_cortthick_qc_uni$paired <- paste(smrip_cortthick_qc_uni$subjectkey, smrip_cortthick_qc_uni$eventname)

# These are major outlier. Impossible values.
smrip_cortthick_qc_uni_outlier <- smrip_cortthick_qc_uni %>% filter(paired %in% smri_outlier$paired)

smrip_cortthick_qc_uni_norm <- smrip_cortthick_qc_uni %>% filter(!(paired %in% smri_outlier$paired))

skim(smrip_cortthick_qc_uni_outlier)
skim(smrip_cortthick_qc_uni_norm)

nrow(smrip_cortthick_qc_uni_outlier) # 317
nrow(smrip_cortthick_qc_uni_norm) # 15658


### Check why there are so many negative values and very small values

```


# See Distributions of truncated data
```{r}
smri_list <- colnames(smrip_cortthick_qc_uni_norm %>% select(contains("smri")))

for (smri in 1:length(smri_list)) {
  hist(smrip_cortthick_qc_uni_norm[, smri_list[smri]], xlab = smri_list[smri])
}

skim(smrip_cortthick_qc_uni_norm)
```
Check smri_vol_cdk_sufrrh to smri_vol_scs_subcroticalgv

# Remove more extreme outliers
```{r}
# Select values less than 0 or greater than 10
more_outliers <- smrip_cortthick_qc_uni_norm %>% 
  filter_at(vars(smri_thick_cdk_banksstslh:smri_thick_cdk_mean), any_vars(. < 0 |. > 10))

smrip_cortthick_qc_uni_big_norm <- anti_join(smrip_cortthick_qc_uni_norm, more_outliers, by = "paired")

# This should be the dataframe with extreme outlier removed.
smrip_cortthick_qc_uni_big_norm

# Remove any value below 100 which would be impossible
#smrip_cortthick_qc_uni_big_norm <- 
#  smrip_cortthick_qc_uni_norm[!apply(smrip_cortthick_qc_uni_norm %>% 
#                                    select("smri_vol_cdk_banksstslh":"smri_vol_scs_subcorticalgv") < 100, 
#                                1, any, na.rm=TRUE), ]

```


# Check Distributions of all outliers removed
```{r}
smri_list <- colnames(smrip_cortthick_qc_uni_big_norm %>% select(contains("smri")))

for (smri in 1:length(smri_list)) {
  hist(smrip_cortthick_qc_uni_big_norm[, smri_list[smri]], xlab = smri_list[smri])
}

```
Seems pretty good. Will ask about large values.

# Remove a few more extreme outliers
```{r}
smrip_cortthick_qc_uni_big_norm <- smrip_cortthick_qc_uni_big_norm %>% filter(!(smri_thick_cdk_fusiformlh < 2 |
                                            smri_thick_cdk_linguallh < 1.5 |
                                            smri_thick_cdk_ptcatelh < 1.5 | 
                                            smri_thick_cdk_sufrlh < 2 | 
                                            smri_thick_cdk_sutmlh < 2 | 
                                            smri_thick_cdk_trvtmlh > 3.5 |
                                            smri_thick_cdk_banksstsrh > 3.5 |
                                            smri_thick_cdk_cdacaterh < 1.5 | 
                                            smri_thick_cdk_cuneusrh > 3 | 
                                            smri_thick_cdk_fusiformrh < 2 | 
                                            smri_thick_cdk_ifplrh > 3.5 | 
                                            smri_thick_cdk_ihcaterh < 1.5 | smri_thick_cdk_ihcaterh > 3.5 | 
                                            smri_thick_cdk_paracnrh > 3.5 | 
                                            smri_thick_cdk_parsobisrh > 4 | 
                                            smri_thick_cdk_periccrh > 2.5 | 
                                            smri_thick_cdk_ptcaterh < 2 | 
                                            smri_thick_cdk_smrh < 2 | 
                                            smri_thick_cf2_ptlh > 3 |
                                            smri_thick_cf4_frlh > 3.5 | 
                                            smri_thick_cf4_occlh > 3.5 | 
                                            smri_thick_cf4_pllh < 2 | smri_thick_cf4_pllh > 3.5 | 
                                            smri_thick_cf12_cnlh > 3.25 | 
                                            smri_thick_cf12_occlh > 3 | 
                                            smri_thick_cf12_ptoltmlh > 3.5 | 
                                            smri_thick_cf4_plrh < 2 | 
                                            smri_thick_cdk_meanlh < 2 | 
                                            smri_thick_cdk_meanrh < 2 | 
                                            smri_thick_cdk_mean < 2 | smri_thick_cdk_mean > 4 | 
                                            smri_thick_cf12_aomtmlh < 2 | 
                                            smri_thick_cf12_parsopclh > 3.5 | 
                                            smri_thick_cf12_suplrh > 3.5 | 
                                            smri_thick_cf12_aomtmrh < 2 | 
                                            smri_thick_cf12_pcrh > 3.5))

smrip_cortthick_qc_uni_big_norm %>% filter((smri_thick_cdk_fusiformlh < 2 |
                                            smri_thick_cdk_linguallh < 1.5 |
                                            smri_thick_cdk_ptcatelh < 1.5 | 
                                            smri_thick_cdk_sufrlh < 2 | 
                                            smri_thick_cdk_sutmlh < 2 | 
                                            smri_thick_cdk_trvtmlh > 3.5 |
                                            smri_thick_cdk_banksstsrh > 3.5 |
                                            smri_thick_cdk_cdacaterh < 1.5 | 
                                            smri_thick_cdk_cuneusrh > 3 | 
                                            smri_thick_cdk_fusiformrh < 2 | 
                                            smri_thick_cdk_ifplrh > 3.5 | 
                                            smri_thick_cdk_ihcaterh < 1.5 | smri_thick_cdk_ihcaterh > 3.5 | 
                                            smri_thick_cdk_paracnrh > 3.5 | 
                                            smri_thick_cdk_parsobisrh > 4 | 
                                            smri_thick_cdk_periccrh > 2.5 | 
                                            smri_thick_cdk_ptcaterh < 2 | 
                                            smri_thick_cdk_smrh < 2 | 
                                            smri_thick_cf2_ptlh > 3 |
                                            smri_thick_cf4_frlh > 3.5 | 
                                            smri_thick_cf4_occlh > 3.5 | 
                                            smri_thick_cf4_pllh < 2 | smri_thick_cf4_pllh > 3.5 | 
                                            smri_thick_cf12_cnlh > 3.25 | 
                                            smri_thick_cf12_occlh > 3 | 
                                            smri_thick_cf12_ptoltmlh > 3.5 | 
                                            smri_thick_cf4_plrh < 2 | 
                                            smri_thick_cdk_meanlh < 2 | 
                                            smri_thick_cdk_meanrh < 2 | 
                                            smri_thick_cdk_mean < 2 | smri_thick_cdk_mean > 4 | 
                                            smri_thick_cf12_aomtmlh < 2 | 
                                            smri_thick_cf12_parsopclh > 3.5 | 
                                            smri_thick_cf12_suplrh > 3.5 | 
                                            smri_thick_cf12_aomtmrh < 2 | 
                                            smri_thick_cf12_pcrh > 3.5))





```


# Check Distributions again
```{r}

smri_list <- colnames(smrip_cortthick_qc_uni_big_norm %>% select(contains("smri")))

for (smri in 1:length(smri_list)) {
  hist(smrip_cortthick_qc_uni_big_norm[, smri_list[smri]], xlab = smri_list[smri])
}
```




# Split into Timepoints
## For comparison with Bio Age, we only need year 2
```{r}
smrip_cortthick_qc_uni_y2 <- smrip_cortthick_qc_uni_big_norm %>% filter(eventname == "2_year_follow_up_y_arm_1")
smrip_cortthick_qc_uni_base <- smrip_cortthick_qc_uni_big_norm %>% filter(eventname == "baseline_year_1_arm_1")

nrow(smrip_cortthick_qc_uni_y2) # 6245
nrow(smrip_cortthick_qc_uni_base) # 9092

write.csv(smrip_cortthick_qc_uni_big_norm, "/Users/hansoochang/Drexel/ABCD/data/smrip_cortthick.csv")
write.csv(smrip_cortthick_qc_uni_y2, "/Users/hansoochang/Drexel/ABCD/data/smrip_cortthick_y2.csv")
write.csv(smrip_cortthick_qc_uni_base, "/Users/hansoochang/Drexel/ABCD/data/smrip_cortthick_base.csv")
```


# Check the relationship between CA and ROIs as a table
## IMPORTANT: Kendall's Tau takes too long in regular cor() function. Use cor.fk from pcaPP package
```{r}
smri_list <- colnames(smrip_cortthick_qc_uni_y2 %>% dplyr::select(contains(c("smri"))))

# Non-Parametric
cor_sub <- smrip_cortthick_qc_uni_y2 %>% dplyr::select(contains(c("smri", "interview_age")))
cor_sub$interview_age <- as.numeric(cor_sub$interview_age)
cor_sub <- cor_sub[complete.cases(cor_sub), ]

cor_matrix <- cor.fk(cor_sub)
correlations <- cor_matrix["interview_age", smri_list]

# Matrix with highest absolute correlations between ROI and age
cor_mat_kendall <- sort(abs(correlations), decreasing = T)

# Convert to Pearson's r and Cohen's d for effect size comparison
cor_mat_kendall_to_r <- sin(pi*0.5*cor_mat_kendall)
cor_mat_kendall_to_d <- 2*cor_mat_kendall_to_r/sqrt(1-cor_mat_kendall_to_r**2)
```

# Subset original df 
# Get any ROIs with r above .1 
```{r}
sub_rois_cortthick <- as.data.frame(cor_mat_kendall_to_r) %>% filter(cor_mat_kendall_to_r < 0.1) %>% select(-cor_mat_kendall_to_r)

sub_rois_cortthick <- tibble::rownames_to_column(sub_rois_cortthick, "a") 
sub_rois_cortthick_list <- sub_rois_cortthick$a
sub_rois_cortthick_list

# From PhysMentHealth.Rmd
unhealthy_cortthick_cbcl_phys <- read.csv("/Users/hansoochang/Drexel/ABCD/data/unhealthy_cortthick_cbcl_phys.csv")
healthy_cortthick_cbcl_phys <- read.csv("/Users/hansoochang/Drexel/ABCD/data/healthy_cortthick_cbcl_phys.csv")

healthy_cortthick_cbcl_phys_sub <- healthy_cortthick_cbcl_phys %>% select(-sub_rois_cortthick_list)
unhealthy_cortthick_cbcl_phys_sub <- unhealthy_cortthick_cbcl_phys %>% select(-sub_rois_cortthick_list)

write.csv(healthy_cortthick_cbcl_phys_sub, "/Users/hansoochang/Drexel/ABCD/data/healthy_cortthick_cbcl_phys_sub.csv")
write.csv(unhealthy_cortthick_cbcl_phys_sub, "/Users/hansoochang/Drexel/ABCD/data/unhealthy_cortthick_cbcl_phys_sub.csv")
```

# Get Scanner Info
```{r}

```













