---
title: "BrainAge_KDM_CortThick"
output: html_document
date: "2023-03-22"
---


```{r}
#install.packages("flexsurv")
#devtools::install_github("dayoonkwon/BioAge")
library(BioAge)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(Metrics)
library(stats)
library(skimr)
```


Since KDM requires chronological age, we need chronological age in our biological markers. Convert age in months to age in years. 


# Split into and test set from healthy cohort
```{r}
healthy_cortthick_cbcl_phys <- read.csv("/Users/hansoochang/Drexel/ABCD/data/healthy_cortthick_cbcl_phys.csv")
unhealthy_cortthick_cbcl_phys <- read.csv("/Users/hansoochang/Drexel/ABCD/data/unhealthy_cortthick_cbcl_phys.csv")


healthy_cortthick_cbcl_phys$age <- as.numeric(healthy_cortthick_cbcl_phys$interview_age)/12
healthy_cortthick_cbcl_phys$gender <- healthy_cortthick_cbcl_phys$sex.x

unhealthy_cortthick_cbcl_phys$age <- as.numeric(unhealthy_cortthick_cbcl_phys$interview_age)/12 
unhealthy_cortthick_cbcl_phys$gender <- unhealthy_cortthick_cbcl_phys$sex.x

# about 4 participants with no age info
healthy_cortthick_cbcl_phys <- healthy_cortthick_cbcl_phys[complete.cases(healthy_cortthick_cbcl_phys), ]
unhealthy_cortthick_cbcl_phys <- unhealthy_cortthick_cbcl_phys[complete.cases(unhealthy_cortthick_cbcl_phys), ]

colnames(healthy_cortthick_cbcl_phys)
nrow(healthy_cortthick_cbcl_phys) # 2767
nrow(unhealthy_cortthick_cbcl_phys) # 1104
```

# Check for anything weird
```{r}
skim(unhealthy_cortthick_cbcl_phys)
skim(healthy_cortthick_cbcl_phys)
```


```{r}
set.seed(18)
#bm_healthy <- merge(bm, healthy_cbcl_phys, by = "subjectkey", all = FALSE)
#nrow(bm_healthy) # 375
#bm_unhealthy <- merge(bm, unhealthy_cbcl, by = "subjectkey", all = FALSE)
#nrow(bm_unhealthy) # 283

### Healthy Training ###
samp <- sample(c(TRUE, FALSE), nrow(healthy_cortthick_cbcl_phys), replace=TRUE, prob=c(0.6,0.4))
cortthick_train <- healthy_cortthick_cbcl_phys[samp, ]
cortthick_test <- healthy_cortthick_cbcl_phys[!samp, ]

nrow(cortthick_train) # 1649
nrow(cortthick_test) # 1118

```


# Train
```{r}
cortthick_markers <- colnames(cortthick_train %>% select(contains("smri")))
# kdm_train_male <- kdm_calc(male_train, biomarkers = markers)
# kdm_train_female <- kdm_calc(female_train, biomarkers = markers)'

kdm_train_cortthick <- kdm_calc(cortthick_train, biomarkers = cortthick_markers)
```


Correlation tests between kdm_age and chronological age
```{r}
cor.test(as.numeric(kdm_train_cortthick$data$kdm), as.numeric(kdm_train_cortthick$data$age))
rmse(as.numeric(kdm_train_cortthick$data$kdm), as.numeric(kdm_train_cortthick$data$age))
mae(as.numeric(kdm_train_cortthick$data$kdm), as.numeric(kdm_train_cortthick$data$age))

# Female
# cor.test(as.numeric(kdm_train_female$data$kdm), as.numeric(kdm_train_female$data$age))
# rmse(as.numeric(kdm_train_female$data$kdm), as.numeric(kdm_train_female$data$age))
# mae(as.numeric(kdm_train_female$data$kdm), as.numeric(kdm_train_female$data$age))


#plot age vs bioage
#plot_ba(kdm_train$data, "kdm", "KDM\nBiological Age")

```

# Project to test
```{r}
# 
kdm_test_cortthick <- kdm_calc(cortthick_test, biomarkers = cortthick_markers, fit = kdm_train_cortthick$fit, s_ba2 = 
                       kdm_train_cortthick$fit$s_ba2)

# Female
# kdm_female <- kdm_calc(female_test, biomarkers = markers, fit = kdm_train_female$fit, s_ba2 = 
#                        kdm_train_female$fit$s_ba2)


# kdm_data <- rbind(kdm_male$data, kdm_female$data)
kdm_data_cortthick <- kdm_test_cortthick$data
```

# Correlation tests between kdm_age (both genders) and chronological age
```{r}
cor.test(as.numeric(kdm_data_cortthick$kdm), as.numeric(kdm_data_cortthick$age))
rmse(as.numeric(kdm_data_cortthick$kdm), as.numeric(kdm_data_cortthick$age))
mae(as.numeric(kdm_data_cortthick$kdm), as.numeric(kdm_data_cortthick$age))


#plot age vs bioage
plot_ba(kdm_data_cortthick, "kdm", "KDM\nBiological Age")

```

# plot kdm_advance
```{r}
#plot age vs bioage
plot_ba(kdm_data_cortthick, "kdm", "KDM\nBiological Age")

plot_ba(kdm_data_cortthick, "kdm_advance", "KDM\nBiological Age")

```


# Project trained model onto Unhealthy cohort
## Get prediction
```{r}
# project to unhealthy Group
# males_unhealthy <- bm_unhealthy[which(bm_unhealthy$gender == "M"),]

kdm_unhealthy_cortthick <- kdm_calc(unhealthy_cortthick_cbcl_phys, biomarkers = cortthick_markers, fit = kdm_train_cortthick$fit, s_ba2 = 
                       kdm_train_cortthick$fit$s_ba2)

# Females
#females_unhealthy <- bm_unhealthy[which(bm_unhealthy$gender == "F"),]

#kdm_female_unhealthy <- kdm_calc(females_unhealthy, biomarkers = markers, fit = #kdm_train_female$fit, s_ba2 = 
#                      kdm_train_female$fit$s_ba2)


# merge
# kdm_data_unhealthy <- rbind(kdm_male_unhealthy$data, kdm_female_unhealthy$data)
kdm_data_unhealthy_cortthick <- kdm_unhealthy_cortthick$data 

```


# Correlation test between KDM age and Chronological Age for Unhealthy cohort
```{r}
cor.test(as.numeric(kdm_data_unhealthy_cortthick$kdm), as.numeric(kdm_data_unhealthy_cortthick$age))
rmse(as.numeric(kdm_data_unhealthy_cortthick$kdm), as.numeric(kdm_data_unhealthy_cortthick$age))
mae(as.numeric(kdm_data_unhealthy_cortthick$kdm), as.numeric(kdm_data_unhealthy_cortthick$age))

```

```{r}
plot_ba(kdm_data_unhealthy_cortthick, "kdm", "KDM\nBiological Age")
plot(x = kdm_data_unhealthy_cortthick$age, y = kdm_data_unhealthy_cortthick$kdm_advance)
```

# Test for Normality
```{r}
# variances are approximately = 1

# KDM Biological Age
shapiro.test(kdm_data_cortthick$kdm)
shapiro.test(kdm_data_unhealthy_cortthick$kdm)

ggdensity(kdm_data_cortthick$kdm)
ggdensity(kdm_data_unhealthy_cortthick$kdm)
ggdensity(kdm_data_cortthick$age) # For some reason, less 12 year olds in the data

ggqqplot(kdm_data_cortthick$kdm)
ggqqplot(kdm_data_unhealthy_cortthick$kdm)
```


```{r}
hist(kdm_data_cortthick$kdm)
hist(kdm_data_cortthick$kdm_advance)

hist(kdm_data_unhealthy_cortthick$kdm)
hist(kdm_data_unhealthy_cortthick$kdm_advance)
```

# Check some outliers
```{r}
kdm_data_cortthick %>% filter(abs(kdm_advance) > 10)
kdm_data_unhealthy_cortthick %>% filter(abs(kdm_advance) > 10)


kdm_data_cortthick %>% filter(smri_thick_cdk_mean < 1)
kdm_data_cortthick %>% filter(kdm > 20)

```



# Check the relationship between CA and ROIs Visually
```{r}
smri_list <- colnames(kdm_data_cortthick %>% select(contains("smri")))
for (smri in smri_list) {
  print(ggscatter(kdm_data_cortthick, x = "age", y = smri, 
          add = "reg.line", conf.int = TRUE, 
          xlab = "age", ylab = smri, color = "gender",
          palette = c(M = "blue", F = "pink")) +
    stat_cor(aes(color = gender)) + stat_cor(label.y.npc = 0.8))
}
```

# Check the relationship between CA and ROIs as a table
```{r}
cor_sub <- kdm_data_cortthick %>% select(contains(c("smri", "age")))
cor_sub <- cor_sub %>% select(-interview_age)

cor_matrix <- cor(cor_sub)
correlations <- cor_matrix["age", smri_list]

# Matrix with highest absolute correlations between ROI and age
sort(abs(correlations), decreasing = T)
```



# T-Test
```{r}

sd(kdm_data_cortthick$kdm)
sd(kdm_data_unhealthy_cortthick$kdm)

sd(kdm_data_cortthick$kdm_advance)
sd(kdm_data_unhealthy_cortthick$kdm_advance)

# T-Test for average healthy KDM and unhealthy KDM
t.test(kdm_data_cortthick$kdm, kdm_data_unhealthy_cortthick$kdm, alternative = "two.sided", var.equal = T)

# T-Test for average healthy KDM_advance and unhealthy KDM_advance
t.test(kdm_data_cortthick$kdm_advance, kdm_data_unhealthy_cortthick$kdm_advance, alternative = "two.sided", var.equal = T)

# t.test between chronological ages
t.test(kdm_data_cortthick$age, kdm_data_unhealthy_cortthick$age, alternative = "two.sided", var.equal = TRUE)

```




















