---
title: "KDM_GMV_ContinuousCBCL_year2"
output: html_document
date: "2023-04-07"
---

# Libraries
```{r}
library(dplyr)
library(ggplot2)
library(ggpubr)
```


# Read in data
```{r}
# Read in data
cbcl <- read.delim("/Users/hansoochang/Drexel/ABCD/data/abcd_cbcls01.txt", sep="", fill = TRUE)

table(cbcl$eventname)
```

# Select only Total Scores and by Collection Year
```{r}
cbcl_t <- cbcl %>% select(contains(c("_t", "subjectkey", "eventname", "interview_age")))
cbcl_t <- cbcl_t[-1, ]

# remove some unnecessary columns
cbcl_t <- subset(cbcl_t, select = -c(`cbcl_scr_syn_thought_m`, `cbcl_scr_syn_thought_nm`
                                       , `cbcl_scr_syn_thought_r`,`cbcl_scr_syn_totprob_m`,
                                       `cbcl_scr_syn_totprob_nm`, `cbcl_scr_syn_totprob_r`))

cbcl_t[, 1:20] <- as.numeric(unlist(cbcl_t[, 1:20]))

cbcl_t_baseline <- cbcl_t[cbcl_t$eventname == "baseline_year_1_arm_1", ]
cbcl_t_1year <- cbcl_t[cbcl_t$eventname == "1_year_follow_up_y_arm_1", ]
cbcl_t_2year <- cbcl_t[cbcl_t$eventname == "2_year_follow_up_y_arm_1", ]
cbcl_t_3year <- cbcl_t[cbcl_t$eventname == "3_year_follow_up_y_arm_1", ]

```


# Add family id
```{r}
FamilyId <- read.csv("/Users/hansoochang/Drexel/ABCD/data/acspsw03.csv")
FamilyId <- FamilyId[-1,]
FamilyId[FamilyId == ""] <- NA

FamilyId_rel_family <- FamilyId %>% select(subjectkey, eventname, rel_family_id)

FamilyId_rel_family <- FamilyId_rel_family[complete.cases(FamilyId_rel_family), ]

cbcl_t_baseline <- merge(cbcl_t_baseline, FamilyId_rel_family, by = c("subjectkey"))
cbcl_t_1year <- merge(cbcl_t_1year, FamilyId_rel_family, by = c("subjectkey"))
cbcl_t_2year <- merge(cbcl_t_2year, FamilyId_rel_family, by = c("subjectkey"))
cbcl_t_3year <- merge(cbcl_t_3year, FamilyId_rel_family, by = c("subjectkey"))


```


# Complete Cases
```{r}
cbcl_t_baseline <- cbcl_t_baseline[complete.cases(cbcl_t_baseline), ]
cbcl_t_1year <- cbcl_t_1year[complete.cases(cbcl_t_1year), ]
cbcl_t_2year <- cbcl_t_2year[complete.cases(cbcl_t_2year), ]
cbcl_t_3year <- cbcl_t_3year[complete.cases(cbcl_t_3year), ]


skim(cbcl_t_2year)

```


# Add column with overall total score
```{r}
colnames(cbcl_t_2year)
cbcl_2year_sum <- cbcl_t_2year %>% 
  rowwise() %>% 
  mutate(sumindex = sum(c_across(c(2:21)), na.rm = T))

cbcl_3year_sum <- cbcl_t_3year %>% 
  rowwise() %>% 
  mutate(sumindex = sum(c_across(c(2:21)), na.rm = T))


```



# Merge with GMV Subset group
## The bm file has binomial cbcl scores, we need to change it back here
```{r}
gmv <- read.csv("/Users/hansoochang/Drexel/ABCD/data/gmv_year2_cs.csv")
gmv_merge <- gmv %>% select(-c("cbcl_scr_syn_anxdep_t":"count"))

colnames(gmv)
colnames(cbcl_2year_sum)

cbcl_2year_sum$eventname <- cbcl_2year_sum$eventname.x
cbcl_2year_sum <- select(-c(eventname.x, eventname.y))

gmv_cbcl_2year_sum <- merge(gmv_merge, cbcl_2year_sum, by = c("subjectkey", "eventname"))
 
gmv_cbcl_3year_sum <- merge(gmv_merge, cbcl_3year_sum, by = c("subjectkey"))

gmv_cbcl_3year_sum$eventname <- gmv_cbcl_3year_sum$eventname.x

gmv_cbcl_3year_sum <- gmv_cbcl_3year_sum %>% select(-c(X.3, X.2, X, X.1, X.x, sex.x, X.y))

write.csv(gmv_cbcl_2year_sum, "/Users/hansoochang/Drexel/ABCD/data/gmv_cbcl_2year_sum")
write.csv(gmv_cbcl_3year_sum, "/Users/hansoochang/Drexel/ABCD/data/gmv_cbcl_3year_sum")

nrow(gmv_cbcl_2year_sum) # 2117
nrow(gmv_cbcl_3year_sum) # 1838
```


# Check Distribution
```{r}
ggdensity(gmv_cbcl_2year_sum$sumindex)
ggdensity(gmv_cbcl_3year_sum$sumindex)

ggdensity(gmv_cbcl_2year_sum$cbcl_scr_syn_totprob_t)
ggdensity(gmv_cbcl_3year_sum$cbcl_scr_syn_totprob_t)
```


### 2 year to 2 year ###
# Plot GMV KDM BA and CBCL scores 
```{r}
colnames(gmv_cbcl_2year_sum)

## Find significant correlations first
cbcl_list <- colnames(gmv_cbcl_2year_sum %>% 
                        select(c("cbcl_scr_syn_anxdep_t" : "cbcl_scr_07_stress_t", "sumindex")))

for (cbcl in cbcl_list) {
  print(ggscatter(gmv_cbcl_2year_sum, x = "kdm", y = cbcl, 
          add = "reg.line", conf.int = TRUE, 
          xlab = "kdm", ylab = cbcl, color = "gender",
          palette = c(M = "blue", F = "pink")) +
    stat_cor(aes(color = gender)) + stat_cor(label.y.npc = 0.8))
}


```

### 2 year to 2 year ###
# Plot GMV KDM BA Advance and CBCL scores 
```{r}
colnames(gmv_cbcl_2year_sum)

## Find significant correlations first
cbcl_list <- colnames(gmv_cbcl_2year_sum %>% 
                        select(c("cbcl_scr_syn_anxdep_t" : "cbcl_scr_07_stress_t", "sumindex")))

for (cbcl in cbcl_list) {
  print(ggscatter(gmv_cbcl_2year_sum, x = "kdm_advance", y = cbcl, 
          add = "reg.line", conf.int = TRUE, 
          xlab = "kdm advance", ylab = cbcl, color = "gender",
          palette = c(M = "blue", F = "pink")) +
    stat_cor(aes(color = gender)) + stat_cor(label.y.npc = 0.8))
}


```


### 2 year to 3 year ###
# Plot GMV KDM BA Advance and CBCL scores 
```{r}
colnames(gmv_cbcl_3year_sum)

## Find significant correlations first
cbcl_list <- colnames(gmv_cbcl_3year_sum %>% 
                        select(c("cbcl_scr_syn_anxdep_t" : "cbcl_scr_07_stress_t", "sumindex")))

for (cbcl in cbcl_list) {
  print(ggscatter(gmv_cbcl_3year_sum, x = "kdm_advance", y = cbcl, 
          add = "reg.line", conf.int = TRUE, 
          xlab = "kdm advance", ylab = cbcl, color = "gender",
          palette = c(M = "blue", F = "pink")) +
    stat_cor(aes(color = gender)) + stat_cor(label.y.npc = 0.8))
}


```

# Calculate Differences in Sex via Regression
		○ Y is cbcl score
		○ X ba gap + gender + interaction of gender and kdm_advance
			§ Interaction shows whether gender moderates kdm_advance
		○ Add multilevel model for site and family
```{r}
gmv_cbcl_2year_sum$sex_dummy <- as.factor(ifelse(gmv_cbcl_2year_sum$gender == "M", 1, 0))
colnames(gmv_cbcl_2year_sum)

sex_lmer <- lmer(data = gmv_cbcl_2year_sum, cbcl_scr_syn_external_t ~ kdm_advance*sex_dummy + 
      (1 + kdm_advance + sex_dummy| site_id_l) + (1 | rel_family_id:), REML = F)

summary(sex_lmer)

```

# Iterate through Tests
```{r}
summary(sex_lmer)$coefficients[,c("Estimate", "Pr(>|t|)")]
summary(sex_lmer)$coefficients

fits <- list()
for (cbcl in cbcl_list) { 
  sex_lmer <- lmer(data = gmv_cbcl_2year_sum, paste(cbcl, "~ kdm_advance*sex_dummy + 
      (1 + kdm_advance + sex_dummy| site_id_l) + (1 | rel_family_id)"), REML = F)
  # print(summary(sex_lmer)$coefficients)
  fits <- rbind(fits, cbcl, summary(sex_lmer)$coefficients[,c("Estimate", "Pr(>|t|)")])
}
fits

```



#cpglmm


















