---
title: "SES"
output: html_document
date: "2023-02-28"
---

```{r}
library(skimr)
library(dplyr)
library(ggplot2)
library(ggpubr)
# detach(package:MASS,unload=TRUE)
```




```{r}
rhds <- read.delim("/Users/hansoochang/Drexel/ABCD/data/abcd_rhds01.txt")
rhds <- rhds[-1, ]
dim(rhds)

# skim(rhds)
```

# merge with available BMs
## since file so large, let's cut it down first  
```{r}
bm <- read.csv("/Users/hansoochang/Drexel/ABCD/data/year2_cs.csv")

rhds_ses <- rhds %>% select(contains(c("subjectkey", "eventname","adi", "svi_tot", "opat", "reshist_addr1_coi_z")))

bm$eventname_bm <- bm$eventname.y.x
rhds_ses$eventname_rhds <- rhds_ses$eventname

merge1 <- merge(bm, rhds_ses, by = (c("subjectkey")))
merge1[merge1 == ""] <- NA


# skim(merge1)

```

# Cut out columns witih many missing datapoints and get true summary scores
```{r}
merge2 <- merge1 %>% select(c(subjectkey, sex.x, eventname_rhds, eventname_bm, kdm_advance,
                              reshist_addr1_adi_wsum:reshist_addr1_adi_perc, 
                              reshist_addr2_adi_wsum:reshist_addr2_adi_perc,
                              reshist_addr3_adi_wsum:reshist_addr3_adi_perc,
                              reshist_addr4_adi_wsum,
                              reshist_addr1_coi_z_coi_nat,
                              reshist_addr1_coi_z_ed_nat:reshist_addr1_coi_z_se_nat,
                              reshist_addr1_svi_tot_20142018:reshist_addr3_svi_tot_20142018,
                              reshist_addr1_opat_kfrpp_avg:reshist_addr1_opat_kfrpp_n, 
                              reshist_addr1_opat_kfrppavg_se,
                              reshist_addr2_opat_kfrpp_avg:reshist_addr2_opat_kfrpp_n,
                              reshist_addr3_opat_kfrpp_avg:reshist_addr3_opat_kfrpp_n))
  
  
```


```{r}
skim(merge2)

merge2_2year <- merge2 %>% filter(eventname_rhds == "2_year_follow_up_y_arm_1")

merge2_2year %>% filter(!is.na(reshist_addr2_adi_wsum))
```

# Averaging some Columns
## the adi tests have multiple measures. Average these. There are only about 20 of these cases.
```{r}
merge2_2year <- merge2_2year %>% 
  mutate_at(vars(reshist_addr1_adi_wsum:reshist_addr3_opat_kfrpp_n), as.numeric)

# wsum
merge2_2year$reshist_addr1_adi_wsum_mean <- 
  rowMeans(select(merge2_2year, contains("_wsum")), na.rm = T)

# perc
merge2_2year$reshist_addr1_adi_perc_mean <- 
  rowMeans(select(merge2_2year, contains("_perc")), na.rm = T)

# reshist_addr1_svi_tot_20142018
merge2_2year$reshist_addr1_svi_tot_20142018_mean <- 
  rowMeans(select(merge2_2year, contains("svi_tot_")), na.rm = T)

# reshist_addr1_opat_kfrpp_avg
merge2_2year$reshist_addr1_opat_kfrpp_avg_mean <- 
  rowMeans(select(merge2_2year, contains("kfrpp_avg")), na.rm = T)

# merge2_2year

```

# Select variables of interest after averaging
```{r}
merge3_2year <- merge2_2year %>% select(c(subjectkey:eventname_bm, kdm_advance,
                          reshist_addr1_coi_z_coi_nat:reshist_addr1_coi_z_se_nat,
                          reshist_addr1_adi_wsum_mean,
                          reshist_addr1_svi_tot_20142018_mean:reshist_addr1_opat_kfrpp_avg_mean))


skim(merge3_2year)
merge3_2year_comp <- merge3_2year[complete.cases(merge3_2year), ]
```


```{r}
# merge3_2year_comp

ggdensity(merge3_2year_comp$reshist_addr1_coi_z_coi_nat)
ggdensity(merge3_2year_comp$reshist_addr1_coi_z_ed_nat)
ggdensity(merge3_2year_comp$reshist_addr1_coi_z_he_nat)
ggdensity(merge3_2year_comp$reshist_addr1_coi_z_se_nat)
ggdensity(merge3_2year_comp$reshist_addr1_adi_wsum_mean)
ggdensity(merge3_2year_comp$reshist_addr1_svi_tot_20142018_mean)
ggdensity(merge3_2year_comp$reshist_addr1_opat_kfrpp_avg_mean)


```



```{r}
## Find significant correlations first
ses_list <- colnames(merge3_2year_comp %>% 
                        select(c("reshist_addr1_coi_z_coi_nat" : "reshist_addr1_opat_kfrpp_avg_mean")))

for (ses in ses_list) {
  print(ggscatter(merge3_2year_comp, x = "kdm_advance", y = ses, 
          add = "reg.line", conf.int = TRUE, 
          xlab = "kdm_advance", ylab = ses, color = "sex.x",
          palette = c(M = "blue", F = "pink")) +
    stat_cor(aes(color = sex.x)) + stat_cor(label.y.npc = 0.8))
}
```


# Does SES predict KDM?
Kdm ~ sex*ses(baseline)

## Get baseline
```{r}
# merge2

merge2_base <- merge2 %>% filter(eventname_rhds == "baseline_year_1_arm_1")

```



```{r}
merge2_base <- merge2_base %>% 
  mutate_at(vars(reshist_addr1_adi_wsum:reshist_addr3_opat_kfrpp_n), as.numeric)

# wsum
merge2_base$reshist_addr1_adi_wsum_mean <- 
  rowMeans(select(merge2_base, contains("_wsum")), na.rm = T)

# perc
merge2_base$reshist_addr1_adi_perc_mean <- 
  rowMeans(select(merge2_base, contains("_perc")), na.rm = T)

# reshist_addr1_svi_tot_20142018
merge2_base$reshist_addr1_svi_tot_20142018_mean <- 
  rowMeans(select(merge2_base, contains("svi_tot_")), na.rm = T)

# reshist_addr1_opat_kfrpp_avg
merge2_base$reshist_addr1_opat_kfrpp_avg_mean <- 
  rowMeans(select(merge2_base, contains("kfrpp_avg")), na.rm = T)

# merge2_base
```


```{r}
merge3_base <- merge2_2year %>% select(c(subjectkey:eventname_bm, kdm_advance,
                          reshist_addr1_coi_z_coi_nat:reshist_addr1_coi_z_se_nat,
                          reshist_addr1_adi_wsum_mean,
                          reshist_addr1_svi_tot_20142018_mean:reshist_addr1_opat_kfrpp_avg_mean))


skim(merge3_base)
merge3_base_comp <- merge3_base[complete.cases(merge3_base), ]
```


Kdm (2year) ~ sex*ses(baseline)
I have to iterate through these
```{r}
merge3_base_list <- colnames(merge3_base_comp %>% 
                              select(c("reshist_addr1_coi_z_coi_nat" : 
                                         "reshist_addr1_opat_kfrpp_avg_mean")))

fits <- list()
for (i in 1:length(merge3_base_list)) { 
  print(i)
  mod <- lm(data = merge3_base_comp, 
            formula = paste("kdm_advance ~",  merge3_base_list[i],  
                            "*sex.x" ))
  fits <- rbind(fits, merge3_base_list[i], summary(mod)$coefficients[,c("Estimate", "Pr(>|t|)")])
}
fits


```


# Graph SES baseline and KDM_advance 2 year
```{r}
## Find significant correlations first
ses_list <- colnames(merge3_base_comp %>% 
                        select(c("reshist_addr1_coi_z_coi_nat" : "reshist_addr1_opat_kfrpp_avg_mean")))

for (ses in ses_list) {
  print(ggscatter(merge3_base_comp, x = "kdm_advance", y = ses, 
          add = "reg.line", conf.int = TRUE, 
          xlab = "kdm_advance", ylab = ses, color = "sex.x",
          palette = c(M = "blue", F = "pink")) +
    stat_cor(aes(color = sex.x)) + stat_cor(label.y.npc = 0.8))
}
```










